Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and attentional function: A study based on the parametric g-formula

### 2024-03-14 output/paper/manuscript.qmd
Re-writing discussion basically.

Preliminary evidence is also available for several phthalate metabolites in relation to cognitive development in childhood. Higher levels of di(2-ethylhexyl) phthalate metabolites (including MEHP, MEHHP, and MEOHP) were associated with lower intelligence scores in children aged 2 to 12 years [@HuangChenSu:2015], lower scores of {{< acr iq >}} and verbal {{< acr iq >}}, more omission errors (a measure of inattention), and higher scores of response time variability (a measure of sustained attention) in 6-year old Korean children [@KimHongShin:2017], and poorer fine motor skills in preadolescent boys [@BalalianWhyattLiu:2019].

Few studies have looked into different classes of non-persistent {{< acr edcs >}} [@ShoaffCoullWeuve:2020 in relation with the Conners Attention Deficit Scale and the Behavior Assessment System for Children; @OhKimKannan:2023 in relation with {{< acr adhd >}} symptoms]. Overall, and consistent with our results, these studies seem to provide further evidence of the adverse effects of several non-persistent {{< acr edcs >}} on neurodevelopment (too broad) in children. While not all these studies have investigated effect modification by sex, it seems that these adverse effects tend to be stronger in males [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @YuDuChiou:2016; @ShoaffCoullWeuve:2020], although this difference might depend on the timing of exposure [@].
===

>>>>

Preliminary evidence is also available for several phthalate metabolites in relation to cognitive development in childhood. Higher levels of di(2-ethylhexyl) phthalate metabolites (including MEHP, MEHHP, and MEOHP) were associated with lower intelligence scores in children aged 2 to 12 years [@HuangChenSu:2015], lower scores of {{< acr iq >}} and verbal intelligence, more omission errors (a measure of inattention), and higher scores of response time variability (a measure of sustained attention) in 6-year old Korean children [@KimHongShin:2017], poorer fine motor skills in preadolescent boys [@BalalianWhyattLiu:2019], and lower intelligence scores in 7-year old children [@VilmandBeckBilenberg:2023]. Further associations were found for MEOHP with lower scores of {{< acr iq >}} [@HuangChenSu:2015] and verbal intelligence in Taiwanese children aged 6 to 12 years [@HuangTsaiChen:2017], and for dibutyl phthalate metabolites (MnBP and MiBP) with impaired verbal intelligence [@HuangTsaiChen:2017]. There is further preliminary evidence that associations between certain phthalate metabolites and cognitive abilities vary by timing of exposure assessment [@LiPapandonatosCalafat:2019].

BPA

Few studies have looked into different classes of non-persistent {{< acr edcs >}}. Shoaff et al. investigated cross-sectional associations between multiple {{< acr edcs >}} and {{< acr adhd >}}-related behaviors in 15-year old adolescents, finding a higher risk of {{< acr adhd >}}-related behavior problems with higher levels of antiandrogenic phthalate metabolites (MEHP, MEHHP, MEOHP, MECPP, MnBP, MiBP, MBzP, monohydroxyisobutyl phthalate (MHiBP), monocarboxyoctyl phthalate (MCOP), monoisononyl phthalate (MNP), and monohydroxybutyl phthalate (MHBP)), especially in boys [@ShoaffCoullWeuve:2020].

Oh et al.
[@OhKimKannan:2023]

Overall, and consistent with our results, these studies seem to provide further evidence of the adverse effects of several non-persistent {{< acr edcs >}} on neurodevelopment (too broad) in children. While not all these studies have investigated effect modification by sex, it seems that these adverse effects tend to be stronger in males [@Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @YuDuChiou:2016; @ShoaffCoullWeuve:2020], although this difference might depend on the timing of exposure [@].
===
